Convective heat transport in stratified atmospheres at low and high Mach number

### 2017-02-17 tex/nusselt_paper.tex
adds Re, modifies abstract

\begin{abstract}
Convection in astrophysical systems is stratified and
often occurs at high Rayleigh number (Ra) and low
Mach number (Ma).
Here we study stratified convection in the context of 
plane-parallel, polytropically stratified atmospheres. 
We hold the density stratification (\nrho) and Prandtl 
number (Pr) constant while varying
Ma and Ra to determine 
the behavior of the Nusselt number (Nu), 
which quantifies the efficiency of convective heat transport.
As Ra increases and $\text{Ma} \rightarrow 1$, a scaling 
of Nu $\propto$ Ra$^{0.45}$ is observed.  
As Ra increases to a regime where Ma $\geq 1$,
this scaling gives way to a weaker Nu $\propto$ Ra$^{0.19}$. 
In the regime of Ma $\ll 1$, a consistent
Nu $\propto$ Ra$^{0.31}$ is retrieved,  reminiscent of the 
Nu $\propto$ Ra$^{2/7}$ seen in \RB convection.
\end{abstract}
\maketitle

%%%%% Body of the paper
\section{Introduction}
\refstepcounter{section}
\label{sec:intro}
Convection is essential to heat transport in the cores of high mass stars,
the envelopes of low mass stars, and the atmospheres of terrestrial and 
jovian planets.  The convective dynamics in these astrophysical objects
are influenced by atmospheric 
stratification, which is small in some systems (e.g. massive star cores)
but spans about
14 density scale heights in the Sun's convective envelope.
Understanding the fundamental
properties of compressible convection in stratified media is essential
to characterizing systems in astrophysics and planetary sciences.  
Numerical constraints have often restricted studies of compressible
convection to moderately high Mach number (Ma), appropriate to regions near 
the Sun's surface.  Few fundamental properties of
low Ma convection, which occurs in the deep solar interior, are known.

@@

The efficiency of convection is quantified by the Nusselt number.  
Nu is well-defined in RB convection
as the total flux normalized by the steady-state conductive flux 
\cite{johnston&doering2009, otero&all2002}.
In stratified convection Nu is more difficult to define, and we use
a modified version of a traditional stratified Nusselt number \cite{graham1975,hurlburt&all1984},
\begin{equation}
\text{Nu} \equiv \frac{\langle F_{\text{conv, z}} + F_{\text{cond, z}} - F_{\text{A}}\rangle}
{\langle F_{\text{cond, z}} - F_{\text{A}}\rangle} 
= 1 + \frac{\langle F_{\text{conv, z}}\rangle}{\langle F_{\text{cond, z}} - F_{\text{A}} \rangle}
\label{eqn:nusselt}
\end{equation}
where $F_{\text{conv, z}}$ and $F_{\text{cond, z}}$ are the 
z-components of $\bm{F}_{\text{conv}}$ and $\bm{F}_{\text{cond}}$,
respectively and $\langle \rangle$ implies a volume average.  
$F_{\text{A}} \equiv -\langle\kappa\rangle \partial_z T_{\text{ad}}$ 
is the adiabatic conductive flux and
$\partial_z T_{\text{ad}} \equiv - g / c_{P}$ 
for an ideal gas in hydrostatic equilibrium.  Here we specify
$\langle \kappa \rangle = \langle (\rho_{1}/\rho_{0})\kappa_0 \rangle$.
At large values of $\epsilon$, $\kappa$ evolves significantly as the
density profile evolves.  At low $\epsilon$, $\langle\kappa\rangle \approx \kappa_0$.

We contend that this is the general form of the Nusselt number
and illustrate this with a few limiting
cases. Convection eliminates
entropy stratification.  Under the Boussinesq approximation,
in which density variations are ignored, entropy stratification 
is directly proportional to temperature stratification
and $\partial_z S \rightarrow 0$ when $\partial_z T \rightarrow 0$.  
The familiar Nu in the RB problem is
therefore retrieved from (\ref{eqn:nusselt}) because $\partial_z T_{\text{ad}} = 0$.  
In the case of polytropic convection,  $g \rightarrow 0$
as $\epsilon \rightarrow m_{\text{ad}} + 1$.
In this limit, $\partial_z T_{\text{ad}} \rightarrow 0$ and the
definition of the RB Nu is appropriate to use \cite{brandenburg&all2005}. 
However, as $\epsilon \rightarrow 0$, 
$\partial_z T_{0}\rightarrow \partial_z T_{\text{ad}}$ and increasingly
smaller velocity and thermodynamic perturbations 
achieve $\partial_z S = 0$ (Fig. \ref{fig:ma_v_eps}a).
These perturbations carry what little flux exists in 
excess of the adiabatic, so the removal of 
$F_{\text{A}}$ in the numerator and denominator of 
(\ref{eqn:nusselt}) is essential for a reasonable Nu measurement.

@@

\subsection{acknowledgements}
EHA acknowledges the support of the University of Colorado's George 
Ellery Hale Graduate Student Fellowship.
This work was additionally supported by  NASA LWS grant number NNX16AC92G.  
Computations were conducted 
with support by the NASA High End Computing (HEC) Program through the NASA 
Advanced Supercomputing (NAS) Division at Ames Research Center on Pleiades
with allocations GID s1647 and GID g26133.
We thank Axel Brandenburg, Keith Julien, Mark Rast, and Jeff Oishi for many useful discussions.

>>>>

\begin{abstract}
Convection in astrophysical systems is stratified and
often occurs at high Rayleigh number (Ra) and low
Mach number (Ma).
Here we study stratified convection in the context of 
plane-parallel, polytropically stratified atmospheres. 
We hold the density stratification ($n_{\rho}$) and Prandtl 
number (Pr) constant while varying the superadiabaticity
and Ra.  We find that Ma is primarily controlled by the
superadiabaticity of the atmosphere.  We also examine
the behavior of the Nusselt number (Nu), 
which quantifies the efficiency of convective heat transport,
and the Reynolds number (Re), which quantifies how turbulent our
solutions are.
%As Ra increases and $\text{Ma} \rightarrow 1$, a scaling 
%of Nu $\propto$ Ra$^{0.45}$ is observed.  
%As Ra increases to a regime where Ma $\geq 1$,
%this scaling gives way to a weaker Nu $\propto$ Ra$^{0.19}$. 
%In the regime of Ma $\ll 1$, a consistent
%Nu $\propto$ Ra$^{0.31}$ is retrieved,  reminiscent of the 
%Nu $\propto$ Ra$^{2/7}$ seen in \RB convection.
\end{abstract}
\maketitle

%%%%% Body of the paper
\section{Introduction}
\refstepcounter{section}
\label{sec:intro}
Convective heat transport is essential
in the cores of high mass stars,
the envelopes of low mass stars, and the atmospheres of terrestrial and 
jovian planets.  These astrophysical systems have density stratifications
ranging from a few density scale heights (e.g. massive star cores)
up to 14 density scale heights in the convective envelopes of low
mass stars such as the Sun.
Atmospheric stratification presumably modifies the fundamental
dynamics and heat transport mechanisms of these atmospheres.
Understanding the fundamental
properties of compressible convection in stratified media is essential
to characterizing systems in astrophysics and planetary sciences.  
Numerical constraints have often restricted studies of compressible
convection to moderately high Mach number (Ma), appropriate to regions near 
the Sun's surface.  Few fundamental properties of
low Ma convection, which occurs in the deep solar interior, are known.

@@

The efficiency of convection is quantified by the Nusselt number.  
Nu is well-defined in RB convection
as the total flux normalized by the steady-state conductive flux 
\cite{johnston&doering2009, otero&all2002}.
In stratified convection Nu is more difficult to define, and we use
a modified version of a traditional stratified Nusselt number \cite{graham1975,hurlburt&all1984},
\begin{equation}
\text{Nu} \equiv \frac{\langle F_{\text{conv, z}} + F_{\text{cond, z}} - F_{\text{A}}\rangle}
{\langle F_{\text{cond, z}} - F_{\text{A}}\rangle} 
= 1 + \frac{\langle F_{\text{conv, z}}\rangle}{\langle F_{\text{cond, z}} - F_{\text{A}} \rangle}
\label{eqn:nusselt}
\end{equation}
where $F_{\text{conv, z}}$ and $F_{\text{cond, z}}$ are the 
z-components of $\bm{F}_{\text{conv}}$ and $\bm{F}_{\text{cond}}$,
respectively and $\langle \rangle$ implies a volume average.  
$F_{\text{A}} \equiv -\langle\kappa\rangle \partial_z T_{\text{ad}}$ 
is the adiabatic conductive flux and
$\partial_z T_{\text{ad}} \equiv - g / c_{P}$ 
for an ideal gas in hydrostatic equilibrium.  Here we specify
$\langle \kappa \rangle = \langle (\rho_{1}/\rho_{0})\kappa_0 \rangle$.
At large values of $\epsilon$, $\kappa$ evolves significantly as the
density profile evolves.  At low $\epsilon$, $\langle\kappa\rangle \approx \kappa_0$.
In the Boussinesq approximation, under which $\grad S = 0$ only when 
$\grad T = 0$, this definition reduces the the traditional definition
of the Nusselt number \cite{otero&all2002, johnston&doering2009}.

The Reynolds number and Peclet number,
\begin{equation}
\text{Re} = \frac{|\bm{u}| L_z}{\nu};\,\,\,\,\text{Pe} = \text{Pr}\,\text{Re},
\end{equation}
quantify the importance of advection to diffusion in the evolved
convective state.  Our choice of $\{\nu,\chi\}\propto \rho_0^{-1}$ drastically changes
the value of Re between the top and bottom of the atmosphere.  We report values of
Re at the midplane ($z=L_z/2$) of the atmosphere.

@@

\subsection{acknowledgements}
EHA acknowledges the support of the University of Colorado's George 
Ellery Hale Graduate Student Fellowship.
This work was additionally supported by  NASA LWS grant number NNX16AC92G.  
Computations were conducted 
with support by the NASA High End Computing (HEC) Program through the NASA 
Advanced Supercomputing (NAS) Division at Ames Research Center on Pleiades
with allocations GID s1647 and GID g26133 (\textbf{AND ALSO THE NEW ONE AS OF 2017}).
We thank Axel Brandenburg, Keith Julien, Mark Rast, and Jeff Oishi for many useful discussions.
